### 2021-05-07 paper/paper.tex
tex

% @Author: UnsignedByte
% @Date:   11:12:32, 08-Dec-2020
% @Last Modified by:   UnsignedByte
% @Last Modified time: 2021-05-06 23:13:10

\documentclass{article}
\usepackage{amsmath}
\usepackage{graphicx}
\graphicspath{ {./images/} }
\usepackage[%  
    colorlinks=true,
    pdfborder={0 0 0},
    linkcolor=black
]{hyperref}
\title{Generalized Derivation for the motion of an n-pendulum system}
\author{Edmund Lam \and Nicolas Wieczorek}
\begin{document}
\begin{titlepage}
	\maketitle
	\thispagestyle{empty}
\end{titlepage}
\newpage
  \tableofcontents
\newpage

\section{Introduction}
All source code for this project is public on \href{https://github.com/UnsignedByte/n-pendulums}{Github}.

\section{Lagrangian Approach}
In contrast to the limitations of the Newtonian Free Body Diagram approach to calculating propogated forces amongst an $n$-pendulum system, the use of Lagrangian Mechanics does not require an analysis of forces. Instead, the total potential and kinetic energy in the system is used to determine equations of motion for each pendulum in the system.
\subsection{Derivation of Equations of Motion for Finite \texorpdfstring{$n$}{n}}
The use of the lagrangian method requires on two main 
\subsubsection{Deriving the Lagrangian \texorpdfstring{$L=T-V$}{L=T-V}}
We begin by defining the kinematic constraints of our problem.
\begin{align*}
	x_i &= x_{i-1} + l\sin{\theta_i}\\
	&= l\sum_{k=1}^i\sin{\theta_k}\\
	y_i &= y_{i-1} - l\cos{\theta_i}\\
	&= -l\sum_{k=1}^i\cos{\theta_k}
\end{align*}
Note that $y_i$ is negative, because $\theta_i$ is to the negative vertical. Next we derive the velocities. These can be calculated simply by taking the derivative of the positions.
\begin{align*}
	\dot{x}_i &= \dot x_{i-1} + \dot\theta_i \cdot l\cos{\theta_i}\\
	&= l\sum_{k=1}^i\dot\theta_k\cos{\theta_k}\\
	\dot{x}_i &= \dot y_{i-1} + \dot\theta_i \cdot l\sin{\theta_i}\\
	&= l\sum_{k=1}^i\dot\theta_k\sin{\theta_k}
\end{align*}
Plugging these in, we can calculate the total potential energy $V=mgh$ and the total kinetic energy $T=mv^2$. First, we solve for $V$ and simplify:
\begin{align*}
	V &= mg\sum_{i=1}^n y_i \\
	&=-mgl\sum_{i=1}^n\sum_{k=1}^i\cos\theta_k
	&\text{(Substitute in $y_i$)}\\
	&=-mgl\sum_{i=1}^n (n-i+1)\cos\theta_i
	&\text{(Each element $\cos\theta_i$ appears $(n-i+1)$ times)}\\
\end{align*}
And we can do the same with $T$.
\begin{align*}
	T &= \frac{m}{2}\sum_{i=1}^n (\dot{x}_i^2+\dot{y}_i^2)\\
	&= \frac{ml^2}{2}\sum_{i=1}^n \left(\left(\sum_{k=1}^i\dot\theta_k\cos{\theta_k}\right)^2+\left(\sum_{k=1}^i\dot\theta_k\sin{\theta_k}\right)^2\right)
	&\text{(Substitute in $\dot x_i$ and $\dot y_i$)}\\
	&= \frac{ml^2}{2}\sum_{i=1}^n \sum_{j=1}^i\sum_{k=1}^i\dot\theta_j\dot\theta_k\left(\cos{\theta_j}\cos{\theta_k}+\sin{\theta_j}\sin{\theta_k}\right)
	&\text{(Expand squared sums)}\\
	&= \frac{ml^2}{2}\sum_{i=1}^n \sum_{j=1}^i\sum_{k=1}^i\dot\theta_j\dot\theta_k\cos(\theta_j-\theta_k)
	&\text{(Simplify using $\cos$ addition formula)}\\
	&= \frac{ml^2}{2} \sum_{j=1}^n\sum_{k=1}^n(n-\max(j,k)+1)\dot\theta_j\dot\theta_k\cos(\theta_j-\theta_k)
	&\text{(Simplify out first sum)}\\
\end{align*}
Next, we can plug $T$ and $V$ into the lagrangian, giving us the following:
\begin{align*}
	L &= T-V \\
	&=  \frac{ml^2}{2} \sum_{j=1}^n\sum_{k=1}^n(n-\max(j,k)+1)\dot\theta_j\dot\theta_k\cos(\theta_j-\theta_k) + mgl\sum_{i=1}^n (n-i+1)\cos\theta_i\\
\end{align*}
\subsubsection{Utilizing Lagrange's Equations of Motion}
Now that we have our $L$ expression, we can use this to solve for the angular acceleration of each pendulum using Lagrange's equations of motion. We start with the equation $0 = \frac{d}{dt}\left(\frac{\partial L}{\partial \dot \theta_i}\right) - \frac{\partial L}{\partial \theta_i}$, and we will use this equation to solve for all $\ddot\theta_i.$

First, we try to solve for $\frac{\partial L}{\partial \dot \theta_i}$ by taking a derivative of $L$ with respect to $\dot\theta_i$. We first make an important observation: all terms in multiple sums of $L$ that do not contain $\dot\theta_i$ can be treated as a constant, and will be removed. We immediately can remove $V,$ which leaves us with the $T$ term. Because $\dot\theta_i$ appears if and only if $j=i$ or $k=i$, we can simplify our equation much further. We can arbitrarily plug in $j=i$ or $k=i$ into our equation. Before this, however, we must separate our sum due to the special case $i=i$. Firstly, the term simplifies to $\dot\theta_i^2$ when $i=i,$ which cannot be differentiated while treating one of the $\dot\theta_i$ as a constant. It is also the only term that would be double counted if we had kept the simple single sum and doubled it. This gives us 3 cases, one for $i=i,$ one for $1\leq k\le i$ and finally for $i+1\leq k \leq n$. The latter two are counted twice, once as $\dot\theta_i\dot\theta_k$ and once as $\dot\theta_k\dot\theta_i,$ thus we need a factor of two before both sums. This gives us the following, which we can further simplify:
\begin{align*}
&\frac{ml^2}{2}\frac{\partial L}{\partial \dot \theta_i}\left((n-i+1)\dot\theta_i^2 + 2\sum_{k=1}^{i-1}(n-i+1)\dot\theta_i\dot\theta_k\cos(\theta_i-\theta_k)\right.\\
&\qquad\qquad+ \left.2\sum_{k=i+1}^n(n-k+1)\dot\theta_i\dot\theta_k\cos(\theta_i-\theta_k)\right)\\
&= ml^2\left((n-i+1)\dot\theta_i + (n-i+1)\sum_{k=1}^{i-1}\dot\theta_k\cos(\theta_i-\theta_k)\right.\\
&\qquad\qquad+ \left.\sum_{k=i+1}^n(n-k+1)\dot\theta_k\cos(\theta_i-\theta_k)\right)
\end{align*}
Next we take the derivative with respect to time:
\begin{align*}
&= ml^2\frac{d}{dt}\left((n-i+1)\dot\theta_i + (n-i+1)\sum_{k=1}^{i-1}\dot\theta_k\cos(\theta_i-\theta_k)\right.\\
&\qquad\qquad+ \left.\sum_{k=i+1}^n(n-k+1)\dot\theta_k\cos(\theta_i-\theta_k)\right)\\
&= ml^2\left((n-i+1)\ddot\theta_i + (n-i+1)\sum_{k=1}^{i-1}\left(\ddot\theta_k\cos(\theta_i-\theta_k)-\dot\theta_k(\dot\theta_i-\dot\theta_k)\sin(\theta_i-\theta_k)\right)\right.\\
&\qquad\qquad+ \left.\sum_{k=i+1}^n(n-k+1)\left(\ddot\theta_k\cos(\theta_i-\theta_k)-\dot\theta_k(\dot\theta_i-\dot\theta_k)\sin(\theta_i-\theta_k)\right)\right)\\
\end{align*}
\subsection{Generalization to an Infinite Rope}
e
\end{document}

### 2021-05-07 paper/paper.tex
math

@@
 % @Last Modified by:   UnsignedByte
-% @Last Modified time: 2021-05-06 23:13:10
+% @Last Modified time: 2021-05-07 00:29:29
 
@@
 \end{align*}
+Now that we have the $\frac{d}{dt}\left(\frac{\partial L}{\partial \dot\theta_i}\right)$ term, we want to get $\frac{\partial L}{\partial \theta_i}$ in order to finish our Lagrangian equations.
+We approach this with the same method: simplifying our equation for $L$ with only the terms consisting of $\theta_i$.
+For $T$, we have the same situation: a term only contains $\theta_i$ when $j=i$ or $k=i$.
+This gives us a similar equation, but one where $\dot\theta_i^2$ can be removed.
+This time, however, we must also consider $V$.
+$V$ is easy to simplify: the only term with $\theta_i$ is simply the $ith$ term of the sum.
+This gives us the following equation:
+\begin{align*}
+&\frac{ml^2}{2}\frac{\partial L}{\partial \theta_i}\left(2\sum_{k=1}^{i-1}(n-i+1)\dot\theta_i\dot\theta_k\cos(\theta_i-\theta_k)\right.\\
+&\qquad\qquad+ \left.2\sum_{k=i+1}^n(n-k+1)\dot\theta_i\dot\theta_k\cos(\theta_i-\theta_k)\right)\\
+&\qquad\qquad+ \frac{\partial L}{\partial \theta_i}\left(mgl(n-i+1)\cos\theta_i\right)\\
+&=-ml^2\left(\sum_{k=1}^{i-1}(n-i+1)\dot\theta_i\dot\theta_k\sin(\theta_i-\theta_k)\right.\\
+&\qquad\qquad+ \left.\sum_{k=i+1}^n(n-k+1)\dot\theta_i\dot\theta_k\sin(\theta_i-\theta_k)\right)\\
+&\qquad\qquad- mgl(n-i+1)\sin\theta_i\\
+\end{align*}
+
+Plugging all of this into Lagrange's equations gives and simplifying, we get our final equation.
+\begin{align*}
+0&=ml^2\left((n-i+1)\ddot\theta_i + (n-i+1)\sum_{k=1}^{i-1}\left(\ddot\theta_k\cos(\theta_i-\theta_k)-\dot\theta_k(\dot\theta_i-\dot\theta_k)\sin(\theta_i-\theta_k)\right)\right.\\
+&\qquad\qquad+ \left.\sum_{k=i+1}^n(n-k+1)\left(\ddot\theta_k\cos(\theta_i-\theta_k)-\dot\theta_k(\dot\theta_i-\dot\theta_k)\sin(\theta_i-\theta_k)\right)\right)\\
+&\qquad\qquad+ml^2\left(\sum_{k=1}^{i-1}(n-i+1)\dot\theta_i\dot\theta_k\sin(\theta_i-\theta_k) + \sum_{k=i+1}^n(n-k+1)\dot\theta_i\dot\theta_k\sin(\theta_i-\theta_k)\right)\\
+&\qquad\qquad+ mgl(n-i+1)\sin\theta_i\\
+&=ml^2\left((n-i+1)\ddot\theta_i\right.\\
+&\qquad\qquad+ (n-i+1)\sum_{k=1}^{i-1}\left(\ddot\theta_k\cos(\theta_i-\theta_k)-\dot\theta_k(\dot\theta_i-\dot\theta_k)\sin(\theta_i-\theta_k)+\dot\theta_i\dot\theta_k\sin(\theta_i-\theta_k)\right)\\
+&\qquad\qquad+ \left.\sum_{k=i+1}^n(n-k+1)\left(\ddot\theta_k\cos(\theta_i-\theta_k)-\dot\theta_k(\dot\theta_i-\dot\theta_k)\sin(\theta_i-\theta_k)+\dot\theta_i\dot\theta_k\sin(\theta_i-\theta_k)\right)\right)\\
+&\qquad\qquad+ mgl(n-i+1)\sin\theta_i\\
+&=ml^2\left((n-i+1)\ddot\theta_i + (n-i+1)\sum_{k=1}^{i-1}\left(\ddot\theta_k\cos(\theta_i-\theta_k)-\dot\theta_k^2\sin(\theta_i-\theta_k)\right)\right.\\
+&\qquad\qquad+ \left.\sum_{k=i+1}^n(n-k+1)\left(\ddot\theta_k\cos(\theta_i-\theta_k)-\dot\theta_k^2\sin(\theta_i-\theta_k)\right)\right)\\
+&\qquad\qquad+ mgl(n-i+1)\sin\theta_i
+\end{align*}
 \subsection{Generalization to an Infinite Rope}

### 2021-05-08 paper/paper.tex
infinite

--- 2021-05-07 paper/paper.tex
@@
 % @Last Modified by:   UnsignedByte
-% @Last Modified time: 2021-05-06 23:13:10
+% @Last Modified time: 2021-05-08 00:39:12
 
@@
 \subsection{Derivation of Equations of Motion for Finite \texorpdfstring{$n$}{n}}
-The use of the lagrangian method requires on two main 
+The use of the lagrangian method requires on two main parts: defining constraints and variables, and using them to create equations of motion.
 \subsubsection{Deriving the Lagrangian \texorpdfstring{$L=T-V$}{L=T-V}}
@@
 \end{align*}
-And we can do the same with $T$.
+And we can do the same with $T$, substituting in $v^2=\sum_{i=1}^n\left(\dot{x}_i^2+\dot{y}_i^2\right)$.
 \begin{align*}
-	T &= \frac{m}{2}\sum_{i=1}^n (\dot{x}_i^2+\dot{y}_i^2)\\
+	T &= \frac{m}{2}\sum_{i=1}^n \left(\dot{x}_i^2+\dot{y}_i^2\right)\\
 	&= \frac{ml^2}{2}\sum_{i=1}^n \left(\left(\sum_{k=1}^i\dot\theta_k\cos{\theta_k}\right)^2+\left(\sum_{k=1}^i\dot\theta_k\sin{\theta_k}\right)^2\right)
@@
 \end{align*}
-\subsubsection{Utilizing Lagrange's Equations of Motion}
-Now that we have our $L$ expression, we can use this to solve for the angular acceleration of each pendulum using Lagrange's equations of motion.
+\subsubsection{Utilizing the Euler-Lagrange Equation}
+Now that we have our $L$ expression, we can use this to solve for the angular acceleration of each pendulum using the Euler-Lagrange Equation.
 We start with the equation $0 = \frac{d}{dt}\left(\frac{\partial L}{\partial \dot \theta_i}\right) - \frac{\partial L}{\partial \theta_i}$, and we will use this equation to solve for all $\ddot\theta_i.$
@@
 \end{align*}
+Because $\cos(\theta_i-\theta_i)=1$ and $\sin(\theta_i-\theta_i)=0$, we can combine these sums back into one, reducing our expression back to
+$$ml^2\sum_{k=1}^n(n-\max(i,k)+1)\left(\ddot\theta_k\cos(\theta_i-\theta_k)-\dot\theta_k(\dot\theta_i-\dot\theta_k)\sin(\theta_i-\theta_k)\right).$$
+Note that the derivative could not be taken from the outset, as that would ignore the special case altogether, as well as the double counting occuring in the other cases.
+
+Now that we have the $\frac{d}{dt}\left(\frac{\partial L}{\partial \dot\theta_i}\right)$ term, we want to get $\frac{\partial L}{\partial \theta_i}$ in order to finish our Euler-Lagrange Equation.
+We approach this with the same method: simplifying our equation for $L$ with only the terms consisting of $\theta_i$.
+For $T$, we have the same situation: a term only contains $\theta_i$ when $j=i$ or $k=i$.
+This gives us a similar equation, but one where $\dot\theta_i^2$ can be removed.
+This time, however, we must also consider $V$.
+$V$ is easy to simplify: the only term with $\theta_i$ is simply the $ith$ term of the sum.
+This gives us the following equation:
+\begin{align*}
+&\frac{ml^2}{2}\frac{\partial L}{\partial \theta_i}\left(2\sum_{k=1}^{i-1}(n-i+1)\dot\theta_i\dot\theta_k\cos(\theta_i-\theta_k)\right.\\
+&\qquad\qquad+ \left.2\sum_{k=i+1}^n(n-k+1)\dot\theta_i\dot\theta_k\cos(\theta_i-\theta_k)\right)\\
+&\qquad\qquad+ \frac{\partial L}{\partial \theta_i}\left(mgl(n-i+1)\cos\theta_i\right)\\
+&=-ml^2\left(\sum_{k=1}^{i-1}(n-i+1)\dot\theta_i\dot\theta_k\sin(\theta_i-\theta_k)\right.\\
+&\qquad\qquad+ \left.\sum_{k=i+1}^n(n-k+1)\dot\theta_i\dot\theta_k\sin(\theta_i-\theta_k)\right)\\
+&\qquad\qquad- mgl(n-i+1)\sin\theta_i\\
+\end{align*}
+Using the same reduction as before ($\sin(0)=0$), we can combine the sums again to get a final equation of
+$$-ml^2\sum_{k=1}^n\left((n-\max(i,k)+1)\dot\theta_i\dot\theta_k\sin(\theta_i-\theta_k)\right) - mgl(n-i+1)\sin\theta_i.$$
+
+Plugging all of this into the Euler-Lagrange equations and simplifying, we get our final equation.
+\begin{align*}
+0&=ml^2\sum_{k=1}^n(n-\max(i,k)+1)\left(\ddot\theta_k\cos(\theta_i-\theta_k)-\dot\theta_k(\dot\theta_i-\dot\theta_k)\sin(\theta_i-\theta_k)\right)\\
+&\qquad\qquad+ ml^2\sum_{k=1}^n\left((n-\max(i,k)+1)\dot\theta_i\dot\theta_k\sin(\theta_i-\theta_k)\right) + mgl(n-i+1)\sin\theta_i\\
+&=ml^2\sum_{k=1}^n(n-\max(i,k)+1)\left(\ddot\theta_k\cos(\theta_i-\theta_k)-\dot\theta_k(\dot\theta_i-\dot\theta_k)\sin(\theta_i-\theta_k)+\dot\theta_i\dot\theta_k\sin(\theta_i-\theta_k)\right)\\
+&\qquad\qquad+ mgl(n-i+1)\sin\theta_i\\
+&=ml^2\sum_{k=1}^n(n-\max(i,k)+1)\left(\ddot\theta_k\cos(\theta_i-\theta_k)+\dot\theta_k^2\sin(\theta_i-\theta_k)\right)\\
+&\qquad\qquad+ mgl(n-i+1)\sin\theta_i\\
+\end{align*}
+Because we can substitute into this equation with all $1\leq i\leq n$, we end up with a system of equations with $n$ variables and $n$ equations, which we can solve to calculate the acceleration of each.
 \subsection{Generalization to an Infinite Rope}
-e
+Now that we have our derivation with a finite rope, we generalize this to an infinite rope, with a constant finite mass and density.
+We define our mass-density of the total rope with $\frac{M}{L}$, such that $M$ is the total mass of the rope, $L$ is the total length of the rope, and $\frac{dm}{dl} = \frac{M}{L}$.
+Otherwise, we also need to define the function $\theta_l=\theta(l),$ which gives us the angle $\theta$ given a length $l$ along the rope.
+While the final Euler-Lagrange equation can be simply converted from the one derived in the finite set, the very similar infinite derivation is also included below.
+\subsubsection{Deriving the Lagrangian \texorpdfstring{$L=T-V$}{L=T-V}}
+Once again, we start with the kinematic constraints.
+\begin{align*}
+	x_i &= L\int_0^i\sin\theta_ldl\\
+	y_i &= -L\int_0^i\cos\theta_ldl
+\end{align*}
+Next, we calculate the velocities simply by taking the derivative of their position.
+\begin{align*}
+	\dot x_i &= L\int_0^i\dot\theta_l\cos\theta_ldl\\
+	\dot y_i &= L\int_0^i\dot\theta_l\sin\theta_ldl
+\end{align*}
+With these new equations, calculating the $T$ and $V$ terms becomes much simpler. Once again, we use the same starting equations, and simplify.
+\begin{align*}
+	V &= mg\int_0^L y_idi \\
+	&=-mgL\int_0^L\int_0^i\cos{\theta_l}dldi
+	&\text{(Substitute in $y_l$)}\\
+	&=-mgL\int_0^L(L-l)\cos{\theta_l}dl
+	&\text{(Reduce away inner integral)}
+\end{align*}
+We can do the same with $T$ as well.
+\begin{align*}
+	T &= \frac{m}{2}\int_0^L\left(\dot{x}_i^2+\dot{y}_i^2\right)di\\
+	&= \frac{mL^2}{2}\int_0^L\left(\left(\int_0^i\dot\theta_l\cos\theta_ldl\right)^2+\left(\int_0^i\dot\theta_l\sin\theta_ldl\right)^2\right)di
+	&\text{(Plug in values for sum)}\\
+	&= \frac{mL^2}{2}\int_0^L\int_0^i\int_0^i\dot\theta_j\dot\theta_k\left(\cos\theta_j\cos\theta_k+\sin\theta_j\sin\theta_k\right)djdkdi
+	&\text{(Simplify integrals)}\\
+	&= \frac{mL^2}{2}\int_0^L\int_0^i\int_0^i\dot\theta_j\dot\theta_k\cos(\theta_j-\theta_k)djdkdi
+	&\text{(Combine integrals)}\\
+	&= \frac{mL^2}{2}\int_0^L\int_0^L(L-\max(j,k))\dot\theta_j\dot\theta_k\cos(\theta_j-\theta_k)djdk
+	&\text{(Reduce outer integral)}\\
+\end{align*}
+Combining the two, we get our new $L$ value.
+\begin{align*}
+	L &= T-V \\
+	&= \frac{mL^2}{2}\int_0^L\int_0^L(L-\max(j,k))\dot\theta_j\dot\theta_k\cos(\theta_j-\theta_k)djdk + mgL\int_0^L(L-l)\cos\theta_ldl\\
+\end{align*}
+\subsubsection{Utilizing Lagrange's Equations of Motion}
+With our new $L$ value, we once again use this to solve Lagrange's equations of motion.
+This time we can ignore $i$, giving us $0 = \frac{d}{dt}\left(\frac{\partial L}{\partial \dot \theta}\right) - \frac{\partial L}{\partial \theta}$, and we will use this to find the function $\ddot\theta$.
+
+Once again, we start by solving for $\frac{\partial L}{\partial \dot \theta}$ by taking a derivative of $L$ with respect to $\dot\theta$.
+We can use the same tricks as before, but this time the $i=i$ case does not need to be specially considered, as the integral is continuous and a single individual value makes an insignificant change to the whole.
+\begin{align*}
+&mL^2\frac{\partial L}{\partial \dot \theta_i}\left(\int_0^L(L-\max(i,k))\dot\theta_i\dot\theta_k\cos(\theta_i-\theta_k)dk\right)\\
+&= mL^2\int_0^L(L-\max(i,k))\dot\theta_k\cos(\theta_i-\theta_k)dk
+\end{align*}
+Again, we take the derivative with respect to time:
+\begin{align*}
+&= mL^2\frac{d}{dt}\int_0^L(L-\max(i,k))\dot\theta_k\cos(\theta_i-\theta_k)dk\\
+&= mL^2\int_0^L(L-\max(i,k))\left(\ddot\theta_k\cos(\theta_i-\theta_k)-\dot\theta_k(\dot\theta_i-\dot\theta_k)\sin(\theta_i-\theta_k)\right)dk
+\end{align*}
+With $\frac{d}{dt}\left(\frac{\partial L}{\partial \dot\theta_i}\right)$ finished, we once again calculate the remaining term $\frac{\partial L}{\partial \theta_i}$.
+\begin{align*}
+&mL^2\frac{\partial L}{\partial \theta_i}\left(\int_0^L(L-\max(i,k))\dot\theta_i\dot\theta_k\cos(\theta_i-\theta_k)dk\right) + \frac{\partial L}{\partial \theta_i}mgL(L-i)\cos\theta_i\\
+&= mL^2\int_0^L(L-\max(i,k))\dot\theta_i\dot\theta_k\sin(\theta_i-\theta_k)dk - mgL(L-i)\sin\theta_i\\
+\end{align*}
+Now, we combine this back into the Euler-Lagrange equation, and simplify.
+\begin{align*}
+0&=mL^2\int_0^L(L-\max(i,k))\left(\ddot\theta_k\cos(\theta_i-\theta_k)-\dot\theta_k(\dot\theta_i-\dot\theta_k)\sin(\theta_i-\theta_k)\right)dk\\
+&\qquad\qquad+ mL^2\int_0^L(L-\max(i,k))\dot\theta_i\dot\theta_k\sin(\theta_i-\theta_k)dk + mgL(L-i)\sin\theta_i\\
+&=mL^2\int_0^L(L-\max(i,k))\left(\ddot\theta_k\cos(\theta_i-\theta_k)-\dot\theta_k(\dot\theta_i-\dot\theta_k)\sin(\theta_i-\theta_k) + \dot\theta_i\dot\theta_k\sin(\theta_i-\theta_k)\right)dk\\
+&\qquad\qquad+ mgL(L-i)\sin\theta_i\\
+&=mL^2\int_0^L(L-\max(i,k))\left(\ddot\theta_k\cos(\theta_i-\theta_k)+\dot\theta_k^2\sin(\theta_i-\theta_k)\right)dk+ mgL(L-i)\sin\theta_i\\
+\end{align*}
 \end{document}

### 2021-05-09 include/pendulum.hpp
lagrangian simulation

@@ -1,58 +1,0 @@
-/*
-* @Author: UnsignedByte
-* @Date:   2021-04-15 13:34:46
-* @Last Modified by:   UnsignedByte
-* @Last Modified time: 2021-04-20 13:41:05
-*/
-
-#pragma once
-#include <SFML/Graphics.hpp>
-#include <cmath>
-
-
-struct Pendulum
-{
-public:
-	Pendulum() = default;
-
-	Pendulum(sf::Vector2<long double> pos): _pos(pos)
-	{
-	}
-
-	Pendulum(Pendulum* parent, long double angle, long double length, long double mass): _parent(parent), _theta(angle), l(length), m(mass)
-	{
-		_pos = sf::Vector2<long double>(std::cos(angle)*l, std::sin(angle)*l)+_parent->pos();
-	}
-
-	friend std::ostream& operator<<(std::ostream&, const Pendulum&);
- 
-	sf::Vector2<long double> pos();
-
-	virtual void render(sf::VertexArray&, int);
-
-	/**
-	 * @param fT force of tension exerted
-	 * @param idx Number of pendulums below this
-	 */
-	virtual void update(sf::Vector2<long double> fT);
-
-	long double m = 100.L; //mass
-	long double _vel = 0.L;
-	long double l; //length
-	sf::Vector2<long double> cartVel = sf::Vector2<long double>(10, 10);
-protected:
-	Pendulum* _parent;
-	sf::Vector2<long double> _pos;
-	sf::Vector2<long double> _fT = sf::Vector2<long double>(0, 0);
-	long double _theta; //angle of pendulum
-};
-
-struct Root : Pendulum
-{
-	Root(sf::Vector2<long double> pos): Pendulum(pos)
-	{
-	}
-
-	void render(sf::VertexArray&, int);
-	void update(sf::Vector2<long double>);
-};

### 2021-05-09 include/utils.hpp

@@ -2,15 +2,13 @@
 * @Author: UnsignedByte
 * @Date:   2021-04-15 16:49:10
 * @Last Modified by:   UnsignedByte
-* @Last Modified time: 2021-04-19 23:29:11
+* @Last Modified time: 2021-05-08 16:51:21
 */
 
 #pragma once
 #include <SFML/System.hpp>
 #include <iostream>
-
-const long double DT = 0.02;
-const long double G = 9.8L; //acceleration due to Gravity
+#include <vector>
 // const float mv = static_cast<float>(2*M_PI); // maximum velocity
 
 long double abs(sf::Vector2<long double> v);
@@ -21,3 +19,5 @@ template <typename T> int sign(T x)
 }
 
 std::ostream& operator<<(std::ostream&, const sf::Vector2<long double>);
+std::ostream& operator<<(std::ostream& os, std::vector<std::vector<long double> > matrix);
+std::ostream& operator<<(std::ostream& os, std::vector<long double> matrix);

### 2021-05-09 src/main.cpp

@@ -2,46 +2,99 @@
 * @Author: UnsignedByte
 * @Date:   2021-04-15 13:21:00
 * @Last Modified by:   UnsignedByte
-* @Last Modified time: 2021-04-20 13:14:15
+* @Last Modified time: 2021-05-08 17:50:23
 */
 
 #include <SFML/Graphics.hpp>
 #include <cmath>
 #include <iostream>
 #include <cstring>
-#include "pendulum.hpp"
+#include <vector>
+#include <algorithm>
+#include "utils.hpp"
 
+const int N = 50;
 const int WIDTH = 800;
 const int HEIGHT = 800;
 const sf::Time frameTime = sf::seconds(1/60.f);
-const sf::Time tickTime = sf::seconds(1/1000.f);
-const int COUNT = 1;
-const long double LEN = HEIGHT/static_cast<long double>(COUNT)/2;
-const long double ANGLE = 3*M_PI/4;
+const sf::Time tickTime = sf::seconds(1/400.f);
+const long double L = HEIGHT/static_cast<long double>(N)/2;
+const long double ANGLE = M_PI/2;
+
+const long double DT = 0.0001L;
+const long double G = 9.8L; //acceleration due to Gravity
+const long double M = 5.L; //total mass of rope
+
+std::vector<long double> theta(N,ANGLE); // angular position
+std::vector<long double> thetad(N,0); // angular velocity
+
+sf::VertexArray pendulums(sf::LineStrip, N+1);
+
+void update() {
+	std::vector<std::vector<long double>> system(N,std::vector<long double>(N+1,0)); // system of equations
+
+	// Generate system of equations using terms from paper
+	for (int i = 0; i < N; i++){
+		system[i][N] = -M*G*L*(N-i)*std::sin(theta[i]); // initial constant term
+		for(int j = 0; j < N; j++){
+			int mj = (N-std::max(i, j));
+			system[i][N] -= mj*thetad[j]*thetad[j]*std::sin(theta[i]-theta[j]); // add constant term
+			system[i][j] = mj*std::cos(theta[i]-theta[j]);
+		}
+	}
+
+	//Gaussian elimination on system of equations
+
+	// std::cout << system << '\n';
+
+	for(int i = 0; i < N; i++){
+		for(int j = i+1; j < N; j++){
[…]

### 2021-05-09 src/pendulum.cpp

@@ -1,62 +1,0 @@
-/*
-* @Author: UnsignedByte
-* @Date:   2021-04-15 13:34:57
-* @Last Modified by:   UnsignedByte
-* @Last Modified time: 2021-04-20 13:46:18
-*/
-#include <iostream>
-#include "pendulum.hpp"
-#include <SFML/Graphics.hpp>
-#include <cmath>
-#include "utils.hpp"
-
-void Pendulum::update(sf::Vector2<long double> fT)
-{
-	// std::cout << "updating " << *this << std::endl;
-	sf::Vector2<long double> disp = _pos-_parent->pos();
-	// std::cout << "updating " << disp << std::endl;
-	_theta = std::atan2(disp.y, disp.x);
-
-	fT+=sf::Vector2<long double>(0, G*m);
-
-	long double fTdisptheta = std::atan2(fT.y, fT.x)-_theta;
-	_vel += (abs(fT)*std::sin(fTdisptheta))/(m*l)*DT;
-	// _vel = std::max(std::min(_vel, mv), -mv);
-	// _vel *= 1-DT*0.01;
-	_fT = sf::Vector2<long double>(std::cos(_theta), std::sin(_theta))*(std::cos(fTdisptheta)*abs(fT)+m*_vel*_vel/l); //update with new tension force
-	_theta += _vel*DT;
-	// std::cout << "became " << *this << std::endl;
-	_pos = _parent->pos()+sf::Vector2<long double>(std::cos(_theta)*l, std::sin(_theta)*l);
-	_parent->update(_fT);
-}
-
-void Pendulum::render(sf::VertexArray& arr, int idx)
-{
-	// std::cout << "rendering " << idx << std::endl;
-	arr[idx].position = sf::Vector2f(_pos);
-	arr[idx].color = sf::Color::White;
-	_parent->render(arr, idx+1);
-}
-
-sf::Vector2<long double> Pendulum::pos()
-{
-	return _pos;
-}
-
-std::ostream& operator<<(std::ostream& os, const Pendulum& p)
-{
-	os << "Pendulum\nPosition: " << p._pos << "\nVelocity: " << p._vel << "\nAcceleration: " << p._fT << "\nAngle: " << p._theta;
-	return os;
-}
-
-void Root::update(sf::Vector2<long double> fT)
-{
-	// std::cout << "updating ROOT " << idx << std::endl;
-}
-
-void Root::render(sf::VertexArray& arr, int idx)
-{
[…]

### 2021-05-09 src/utils.cpp

@@ -2,12 +2,13 @@
 * @Author: UnsignedByte
 * @Date:   2021-04-15 16:49:16
 * @Last Modified by:   UnsignedByte
-* @Last Modified time: 2021-04-19 18:51:43
+* @Last Modified time: 2021-05-08 16:50:43
 */
 
 #include <SFML/System.hpp>
 #include <cmath>
-
+#include <iostream>
+#include <vector>
 
 long double abs(sf::Vector2<long double> v)
 {
@@ -18,4 +19,23 @@ std::ostream& operator<<(std::ostream& os, const sf::Vector2<long double> v)
 {
 	os << "("<<v.x<<", "<<v.y<<")";
 	return os;
+}
+
+std::ostream& operator<<(std::ostream& os, std::vector<std::vector<long double> > matrix)
+{
+	for(int i = 0; i < matrix.size(); i++){
+		for(int j = 0; j < matrix[i].size(); j++){
+			os << std::to_string(matrix[i][j]).substr(0,8) << " ";
+		}
+		os << "\b\n";
+	}
+	return os << '\b';
+}
+
+std::ostream& operator<<(std::ostream& os, std::vector<long double> matrix)
+{
+	for(int i = 0; i < matrix.size(); i++){
+		os << std::to_string(matrix[i]).substr(0,8) << " ";
+	}
+	return os << '\b';
 }

### 2021-05-12 paper/paper.tex
lagrangian simulation
video gen

@@
 % @Last Modified by:   UnsignedByte
-% @Last Modified time: 2021-05-08 00:39:12
+% @Last Modified time: 2021-05-11 13:44:56
 
@@
 ]{hyperref}
-\title{Generalized Derivation for the motion of an n-pendulum system}
+\title{Generalized Derivation for the Motion of an n-Pendulum System}
 \author{Edmund Lam \and Nicolas Wieczorek}
@@
 \end{align*}
-Plugging these in, we can calculate the total potential energy $V=mgh$ and the total kinetic energy $T=mv^2$. First, we solve for $V$ and simplify:
+Plugging these in, we can calculate the total potential energy $V=mgh$ and the total kinetic energy $T=\frac{1}{2}mv^2$. First, we solve for $V$ and simplify:
 \begin{align*}
@@
 \end{align*}
+We have our equation 
+$$0=mL^2\int_0^L(L-\max(i,k))\left(\ddot\theta_k\cos(\theta_i-\theta_k)+\dot\theta_k^2\sin(\theta_i-\theta_k)\right)dk+ mgL(L-i)\sin\theta_i$$
+for all $i \in [0,L]$, which we next need to evaluate for all $\ddot\theta_i$.
+\subsubsection{Finding \texorpdfstring{$\ddot\theta_i$}{Angular Acceleration}}
+e
 \end{document}

### 2021-05-12 include/utils.hpp

@@ -2,22 +2,22 @@
 * @Author: UnsignedByte
 * @Date:   2021-04-15 16:49:10
 * @Last Modified by:   UnsignedByte
-* @Last Modified time: 2021-05-08 16:51:21
+* @Last Modified time: 2021-05-08 17:56:15
 */
 
 #pragma once
-#include <SFML/System.hpp>
+// #include <SFML/System.hpp>
 #include <iostream>
 #include <vector>
 // const float mv = static_cast<float>(2*M_PI); // maximum velocity
 
-long double abs(sf::Vector2<long double> v);
+// long double abs(sf::Vector2<long double> v);
 
-template <typename T> int sign(T x)
-{
-	return (0 < x) - (x < 0);
-}
+// template <typename T> int sign(T x)
+// {
+// 	return (0 < x) - (x < 0);
+// }
 
-std::ostream& operator<<(std::ostream&, const sf::Vector2<long double>);
+// std::ostream& operator<<(std::ostream&, const sf::Vector2<long double>);
 std::ostream& operator<<(std::ostream& os, std::vector<std::vector<long double> > matrix);
 std::ostream& operator<<(std::ostream& os, std::vector<long double> matrix);

### 2021-05-12 src/main.cpp

@@ -2,26 +2,43 @@
 * @Author: UnsignedByte
 * @Date:   2021-04-15 13:21:00
 * @Last Modified by:   UnsignedByte
-* @Last Modified time: 2021-05-08 17:50:23
+* @Last Modified time: 2021-05-12 11:36:58
 */
 
 #include <SFML/Graphics.hpp>
+// #include <GLFW/glfw3.h>
 #include <cmath>
 #include <iostream>
 #include <cstring>
 #include <vector>
 #include <algorithm>
+#include <cstdio>
+extern "C"
+{
+	#include <libavcodec/avcodec.h>
+	#include <libavformat/avformat.h>
+	#include <libavutil/avutil.h>
+	#include <libavutil/avio.h>
+}
 #include "utils.hpp"
 
-const int N = 50;
+#define EPSILON 1e-9L
+#undef av_err2str
+#define av_err2str(errnum) \
+av_make_error_string((char*)__builtin_alloca(AV_ERROR_MAX_STRING_SIZE), AV_ERROR_MAX_STRING_SIZE, errnum)
+
+const int N = 2;
 const int WIDTH = 800;
 const int HEIGHT = 800;
-const sf::Time frameTime = sf::seconds(1/60.f);
-const sf::Time tickTime = sf::seconds(1/400.f);
+const float speed = 0.05; // time speed
+const int fps = 30; // fps
+const int duration = 60; // duration in seconds
+// const sf::Time frameTime = sf::seconds(1/60.f);
+// const sf::Time tickTime = sf::seconds(1/400.f);
 const long double L = HEIGHT/static_cast<long double>(N)/2;
-const long double ANGLE = M_PI/2;
+const long double ANGLE = 3*M_PI/4;
 
-const long double DT = 0.0001L;
+const long double DT = 0.00001L; // 1 step of time
 const long double G = 9.8L; //acceleration due to Gravity
 const long double M = 5.L; //total mass of rope
 
@@ -29,8 +46,17 @@ std::vector<long double> theta(N,ANGLE); // angular position
 std::vector<long double> thetad(N,0); // angular velocity
 
 sf::VertexArray pendulums(sf::LineStrip, N+1);
+sf::RenderTexture frame_render;
 
-void update() {
+static void rgbyuv(sf::Color rgb, uint8_t* Y, uint8_t* U, uint8_t* V)
+{
[…]

### 2021-05-12 src/utils.cpp

@@ -2,24 +2,24 @@
 * @Author: UnsignedByte
 * @Date:   2021-04-15 16:49:16
 * @Last Modified by:   UnsignedByte
-* @Last Modified time: 2021-05-08 16:50:43
+* @Last Modified time: 2021-05-08 17:56:18
 */
 
-#include <SFML/System.hpp>
+// #include <SFML/System.hpp>
 #include <cmath>
 #include <iostream>
 #include <vector>
 
-long double abs(sf::Vector2<long double> v)
-{
-	return std::sqrt(v.x*v.x+v.y*v.y);
-}
+// long double abs(sf::Vector2<long double> v)
+// {
+// 	return std::sqrt(v.x*v.x+v.y*v.y);
+// }
 
-std::ostream& operator<<(std::ostream& os, const sf::Vector2<long double> v)
-{
-	os << "("<<v.x<<", "<<v.y<<")";
-	return os;
-}
+// std::ostream& operator<<(std::ostream& os, const sf::Vector2<long double> v)
+// {
+// 	os << "("<<v.x<<", "<<v.y<<")";
+// 	return os;
+// }
 
 std::ostream& operator<<(std::ostream& os, std::vector<std::vector<long double> > matrix)
 {

### 2021-05-13 src/main.cpp
video ouptut fix

@@ -2,7 +2,7 @@
 * @Author: UnsignedByte
 * @Date:   2021-04-15 13:21:00
 * @Last Modified by:   UnsignedByte
-* @Last Modified time: 2021-05-12 11:36:58
+* @Last Modified time: 2021-05-13 00:13:27
 */
 
 #include <SFML/Graphics.hpp>
@@ -18,7 +18,7 @@ extern "C"
 	#include <libavcodec/avcodec.h>
 	#include <libavformat/avformat.h>
 	#include <libavutil/avutil.h>
-	#include <libavutil/avio.h>
+	#include <libavformat/avio.h>
 }
 #include "utils.hpp"
 
@@ -104,7 +104,16 @@ static void update() {
 	// std::cout << "theta: "<< theta << '\n';
 }
 
-static void encode(AVCodecContext* av_codec_context, AVFrame* frame, AVPacket* packet, FILE* file)
+static int write_frame(AVFormatContext* av_format_context, const AVRational& time_base, AVStream *av_stream, AVPacket *packet)
+{
+	/* rescale output packet timestamp values from codec to stream timebase */
+	av_packet_rescale_ts(packet, time_base, av_stream->time_base);
+	packet->stream_index = av_stream->index;
+	// printf("Stream Index: %d, PTS: %" PRId64 ", DTS: %" PRId64 ", pos: %" PRId64 ", duration: %" PRId64 "\n", packet->stream_index, packet->pts, packet->dts, packet->pos, packet->duration);
+	return av_interleaved_write_frame(av_format_context, packet);
+}
+
+static void encode(AVFormatContext* av_format_context, AVCodecContext* av_codec_context, AVStream* av_stream, AVFrame* frame, AVPacket* packet)
 {
 	int res;
 
@@ -114,20 +123,28 @@ static void encode(AVCodecContext* av_codec_context, AVFrame* frame, AVPacket* p
 		exit(1);
 	}
 
-	while(res >= 0) {
+	while(1) {
 		res = avcodec_receive_packet(av_codec_context, packet);
 
 		// all packets sent
 		if (res == AVERROR(EAGAIN) || res == AVERROR_EOF)
 		{
-			return;
+			// av_packet_free(&packet);
+			break;
 		} else if (res < 0) {
 			fprintf(stderr, "Error during encoding.\n");
 			exit(1);
 		}
 
[…]

### 2021-05-13 src/main.cpp
multithread_test

@@ -2,7 +2,7 @@
 * @Author: UnsignedByte
 * @Date:   2021-04-15 13:21:00
 * @Last Modified by:   UnsignedByte
-* @Last Modified time: 2021-05-13 00:13:27
+* @Last Modified time: 2021-05-13 14:32:52
 */
 
 #include <SFML/Graphics.hpp>
@@ -13,6 +13,8 @@
 #include <vector>
 #include <algorithm>
 #include <cstdio>
+#include <pthread.h>
+#include <assert.h>
 extern "C"
 {
 	#include <libavcodec/avcodec.h>
@@ -27,26 +29,23 @@ extern "C"
 #define av_err2str(errnum) \
 av_make_error_string((char*)__builtin_alloca(AV_ERROR_MAX_STRING_SIZE), AV_ERROR_MAX_STRING_SIZE, errnum)
 
-const int N = 2;
-const int WIDTH = 800;
-const int HEIGHT = 800;
-const float speed = 0.05; // time speed
-const int fps = 30; // fps
-const int duration = 60; // duration in seconds
+int N = 2;
+int WIDTH = 800;
+int HEIGHT = 800;
+float speed = 0.05; // time speed
+int fps = 30; // fps
+int duration = 60; // duration in seconds
 // const sf::Time frameTime = sf::seconds(1/60.f);
 // const sf::Time tickTime = sf::seconds(1/400.f);
-const long double L = HEIGHT/static_cast<long double>(N)/2;
+long double L;
 const long double ANGLE = 3*M_PI/4;
 
 const long double DT = 0.00001L; // 1 step of time
 const long double G = 9.8L; //acceleration due to Gravity
 const long double M = 5.L; //total mass of rope
 
-std::vector<long double> theta(N,ANGLE); // angular position
-std::vector<long double> thetad(N,0); // angular velocity
-
-sf::VertexArray pendulums(sf::LineStrip, N+1);
-sf::RenderTexture frame_render;
+std::vector<long double> theta;
+std::vector<long double> thetad;
 
 static void rgbyuv(sf::Color rgb, uint8_t* Y, uint8_t* U, uint8_t* V)
 {
@@ -55,43 +54,77 @@ static void rgbyuv(sf::Color rgb, uint8_t* Y, uint8_t* U, uint8_t* V)
 	*V = static_cast<uint8_t>(rgb.r *  .500000 + rgb.g * -.418688 + rgb.b * -.081312 + 128);
 }
 
[…]

### 2021-05-13 include/utils.hpp
power to LAPACK

@@ -2,7 +2,7 @@
 * @Author: UnsignedByte
 * @Date:   2021-04-15 16:49:10
 * @Last Modified by:   UnsignedByte
-* @Last Modified time: 2021-05-08 17:56:15
+* @Last Modified time: 2021-05-13 15:31:07
 */
 
 #pragma once
@@ -21,3 +21,6 @@
 // std::ostream& operator<<(std::ostream&, const sf::Vector2<long double>);
 std::ostream& operator<<(std::ostream& os, std::vector<std::vector<long double> > matrix);
 std::ostream& operator<<(std::ostream& os, std::vector<long double> matrix);
+std::ostream& operator<<(std::ostream& os, std::vector<std::vector<double> > matrix);
+std::ostream& operator<<(std::ostream& os, std::vector<double> matrix);
+std::ostream& operator<<(std::ostream& os, std::vector<int> matrix);

### 2021-05-13 src/main.cpp

@@ -2,7 +2,7 @@
 * @Author: UnsignedByte
 * @Date:   2021-04-15 13:21:00
 * @Last Modified by:   UnsignedByte
-* @Last Modified time: 2021-05-13 14:32:52
+* @Last Modified time: 2021-05-13 15:33:16
 */
 
 #include <SFML/Graphics.hpp>
@@ -13,14 +13,14 @@
 #include <vector>
 #include <algorithm>
 #include <cstdio>
-#include <pthread.h>
-#include <assert.h>
 extern "C"
 {
 	#include <libavcodec/avcodec.h>
 	#include <libavformat/avformat.h>
 	#include <libavutil/avutil.h>
 	#include <libavformat/avio.h>
+
+	void dgesv_(int *n, int *nrhs, double *a, int *lda, int *ipiv, double *b, int *ldb, int *info);
 }
 #include "utils.hpp"
 
@@ -37,15 +37,15 @@ int fps = 30; // fps
 int duration = 60; // duration in seconds
 // const sf::Time frameTime = sf::seconds(1/60.f);
 // const sf::Time tickTime = sf::seconds(1/400.f);
-long double L;
-const long double ANGLE = 3*M_PI/4;
+double L;
+const double ANGLE = 3*M_PI/4;
 
-const long double DT = 0.00001L; // 1 step of time
-const long double G = 9.8L; //acceleration due to Gravity
-const long double M = 5.L; //total mass of rope
+const double DT = 0.00001L; // 1 step of time
+const double G = 9.8L; //acceleration due to Gravity
+const double M = 5.L; //total mass of rope
 
-std::vector<long double> theta;
-std::vector<long double> thetad;
+std::vector<double> theta;
+std::vector<double> thetad;
 
 static void rgbyuv(sf::Color rgb, uint8_t* Y, uint8_t* U, uint8_t* V)
 {
@@ -54,35 +54,19 @@ static void rgbyuv(sf::Color rgb, uint8_t* Y, uint8_t* U, uint8_t* V)
 	*V = static_cast<uint8_t>(rgb.r *  .500000 + rgb.g * -.418688 + rgb.b * -.081312 + 128);
 }
 
-std::vector<std::vector<long double>> system_eqs; // system_eqs of equations
-
-struct arg {
-	int i;
-	int j;
-	long double factor;
-};
-
-pthread_t* threads;
-arg* thread_args;
-
[…]

### 2021-05-13 src/utils.cpp

@@ -2,7 +2,7 @@
 * @Author: UnsignedByte
 * @Date:   2021-04-15 16:49:16
 * @Last Modified by:   UnsignedByte
-* @Last Modified time: 2021-05-08 17:56:18
+* @Last Modified time: 2021-05-13 15:31:18
 */
 
 // #include <SFML/System.hpp>
@@ -36,6 +36,33 @@ std::ostream& operator<<(std::ostream& os, std::vector<long double> matrix)
 {
 	for(int i = 0; i < matrix.size(); i++){
 		os << std::to_string(matrix[i]).substr(0,8) << " ";
+	}
+	return os << '\b';
+}
+
+std::ostream& operator<<(std::ostream& os, std::vector<std::vector<double> > matrix)
+{
+	for(int i = 0; i < matrix.size(); i++){
+		for(int j = 0; j < matrix[i].size(); j++){
+			os << std::to_string(matrix[i][j]).substr(0,8) << " ";
+		}
+		os << "\b\n";
+	}
+	return os << '\b';
+}
+
+std::ostream& operator<<(std::ostream& os, std::vector<double> matrix)
+{
+	for(int i = 0; i < matrix.size(); i++){
+		os << std::to_string(matrix[i]).substr(0,8) << " ";
+	}
+	return os << '\b';
+}
+
+std::ostream& operator<<(std::ostream& os, std::vector<int> matrix)
+{
+	for(int i = 0; i < matrix.size(); i++){
+		os << matrix[i] << " ";
 	}
 	return os << '\b';
 }

### 2021-05-14 include/utils.hpp
add trail options

@@ -2,11 +2,11 @@
 * @Author: UnsignedByte
 * @Date:   2021-04-15 16:49:10
 * @Last Modified by:   UnsignedByte
-* @Last Modified time: 2021-05-13 15:31:07
+* @Last Modified time: 2021-05-13 17:56:09
 */
 
 #pragma once
-// #include <SFML/System.hpp>
+#include <SFML/Graphics.hpp>
 #include <iostream>
 #include <vector>
 // const float mv = static_cast<float>(2*M_PI); // maximum velocity
@@ -24,3 +24,6 @@ std::ostream& operator<<(std::ostream& os, std::vector<long double> matrix);
 std::ostream& operator<<(std::ostream& os, std::vector<std::vector<double> > matrix);
 std::ostream& operator<<(std::ostream& os, std::vector<double> matrix);
 std::ostream& operator<<(std::ostream& os, std::vector<int> matrix);
+
+sf::Color operator*(sf::Color&c, double n);
+sf::Color& operator*=(sf::Color&c, double n);

### 2021-05-14 src/main.cpp

@@ -2,7 +2,7 @@
 * @Author: UnsignedByte
 * @Date:   2021-04-15 13:21:00
 * @Last Modified by:   UnsignedByte
-* @Last Modified time: 2021-05-13 15:33:16
+* @Last Modified time: 2021-05-13 18:16:11
 */
 
 #include <SFML/Graphics.hpp>
@@ -13,6 +13,7 @@
 #include <vector>
 #include <algorithm>
 #include <cstdio>
+#include <list>
 extern "C"
 {
 	#include <libavcodec/avcodec.h>
@@ -35,8 +36,11 @@ int HEIGHT = 800;
 float speed = 0.05; // time speed
 int fps = 30; // fps
 int duration = 60; // duration in seconds
-// const sf::Time frameTime = sf::seconds(1/60.f);
-// const sf::Time tickTime = sf::seconds(1/400.f);
+float trail_length = 5; //# of seconds of trail to save (will round)
+int _r_trail_length;
+std::vector<int> trail_ids; // pendulum index for each trail
+std::list<sf::Vector2f> trails; // pendulum index for each trail
+
 double L;
 const double ANGLE = 3*M_PI/4;
 
@@ -167,6 +171,32 @@ static int parse_cli(int argc, char** argv, std::vector<char*>& names, std::vect
 	return argi+1;
 }
 
+// https://stackoverflow.com/questions/40629345/fill-array-dynamicly-with-gradient-color-c
+//input: ratio is between 0 to 1
+//output: rgb color
+sf::Color rainbowrgb(double ratio)
+{
+    //we want to normalize ratio so that it fits in to 6 regions
+    //where each region is 256 units long
+    int normalized = int(ratio * 256 * 6);
+
+    //find the distance to the start of the closest region
+    int x = normalized % 256;
+
+    int red = 0, grn = 0, blu = 0;
+    switch(normalized / 256)
+    {
+    case 0: red = 255;      grn = x;        blu = 0;       break;//red
+    case 1: red = 255 - x;  grn = 255;      blu = 0;       break;//yellow
+    case 2: red = 0;        grn = 255;      blu = x;       break;//green
+    case 3: red = 0;        grn = 255 - x;  blu = 255;     break;//cyan
+    case 4: red = x;        grn = 0;        blu = 255;     break;//blue
[…]

### 2021-05-14 src/utils.cpp

@@ -2,10 +2,10 @@
 * @Author: UnsignedByte
 * @Date:   2021-04-15 16:49:16
 * @Last Modified by:   UnsignedByte
-* @Last Modified time: 2021-05-13 15:31:18
+* @Last Modified time: 2021-05-13 18:07:58
 */
 
-// #include <SFML/System.hpp>
+#include <SFML/Graphics.hpp>
 #include <cmath>
 #include <iostream>
 #include <vector>
@@ -65,4 +65,18 @@ std::ostream& operator<<(std::ostream& os, std::vector<int> matrix)
 		os << matrix[i] << " ";
 	}
 	return os << '\b';
+}
+
+sf::Color operator*(sf::Color&c, double n)
+{
+	return sf::Color((int)n*c.r, (int)n*c.g, (int)n*c.b);
+}
+
+sf::Color& operator*=(sf::Color&c, double n)
+{
+	// n = 0.1;
+	c.r*=n;
+	c.g*=n;
+	c.b*=n;
+	return c;
 }

### 2021-05-22 paper/paper.tex
add diagram

@@
 % @Last Modified by:   UnsignedByte
-% @Last Modified time: 2021-05-11 13:44:56
+% @Last Modified time: 2021-05-21 22:45:03
 
@@
 \usepackage{graphicx}
+\usepackage{asymptote}
 \graphicspath{ {./images/} }
@@
 
+\begin{asydef}
+import astylib;
+\end{asydef}
+
 \section{Introduction}
 All source code for this project is public on \href{https://github.com/UnsignedByte/n-pendulums}{Github}.
+
+A single pendulum, is defined by a point of mass suspended by a rigid rod allowed to rotate freely on a joint at the other end of the rod.
+The mechanics of such a pendulum is simple to calculate, even with Newtonian mechanics.
+When attaching a second pendulum to the mass of the first pendulum, we get a system widely known as the "double pendulum." The double pendulum system is widely known to be a chaotic system, meaning a small difference in initial conditions results in widely different motion after a period of time.
+Figure \ref{fig:1} is an example of such a system.
+This method can be continued, adding more and more pendulums, creating a "pendulum chain"--an N pendulum system.
+
+To simplify the problem, we define our system while leaving a number of things constant.
+Our pendulums will have massless rods of constant length $l$, with a mass $m$ at the end of each rod, or at the "joint" between two pendulums.
+Each pendulum will be defined by a single angle $\theta_i$, the angle of the pendulum rod from the vertical.
+
+\begin{figure}[h]
+	\caption{Example double pendulum system with labeled components.}
+	\label{fig:1}
+	\begin{center}
+		\begin{asy}
+		pair R,A,B,AY,BY;
+		real TA,TB,L;
+		L = 100; // length
+		astyscale = 15;
+
+		R = (0,0);
+		TA = 45;
+		A = R+dir(270+TA)*L;
+		AY = (R.x, A.y);
+		TB = -30;
+		B = A+dir(270+TB)*L;
+		BY = (A.x, B.y);
+
+		pen guide = linetype(new real[] {0,5});
+
+		draw(R--A--B);
+		draw(astyangle(AY, R, A));
+		draw(astyangle(B, A, BY));
+		draw(astycomponents(R, A), guide);
+		draw(astycomponents(A, B), guide);
+
+		label("$R$", R, -dir(270+TA));
+		label("$A$", A, E);
+		label("$B$", B, dir(270+TB));
+		label("$l$", R--A, dir(TA));
+		label("$l$", A--B, -dir(TB));
+		label("$y_1$", (R.x, A.y)--R, W);
+		label("$x_1$", (R.x, A.y)--A, S);
+		label("$y_2$", (A.x, B.y)--A, E);
+		label("$x_2$", (A.x, B.y)--B, S);
+		label("$\theta_1$", R+dir(270+TA/2)*astyscale, dir(270+TA/2));
+		label("$\theta_2$", A+dir(270+TB/2)*astyscale, dir(270+TB/2));
+		\end{asy}
+	\end{center}
+\end{figure}
 
@@
 \begin{align*}
-	\dot{x}_i &= \dot x_{i-1} + \dot\theta_i \cdot l\cos{\theta_i}\\
+	\dot x_i &= \dot x_{i-1} + \dot\theta_i \cdot l\cos{\theta_i}\\
 	&= l\sum_{k=1}^i\dot\theta_k\cos{\theta_k}\\
-	\dot{x}_i &= \dot y_{i-1} + \dot\theta_i \cdot l\sin{\theta_i}\\
+	\dot y_i &= \dot y_{i-1} + \dot\theta_i \cdot l\sin{\theta_i}\\
 	&= l\sum_{k=1}^i\dot\theta_k\sin{\theta_k}
@@
 for all $i \in [0,L]$, which we next need to evaluate for all $\ddot\theta_i$.
-\subsubsection{Finding \texorpdfstring{$\ddot\theta_i$}{Angular Acceleration}}
-e
 \end{document}

### 2021-05-22 paper/build.sh

@@ -1,0 +1,19 @@
+#!/bin/bash
+
+logFile="build.log"
+
+touch "$logFile"
+
+echo "Initial Build"
+pdflatex paper.tex > "$logFile"
+
+echo "Building ASY"
+for f in ./*.asy; do
+	echo "Building $f"
+	asy "$f" > "$logFile"
+done
+
+echo "Building refs"
+pdflatex paper.tex > "$logFile"
+echo "Final Build"
+pdflatex paper.tex > "$logFile"

### 2021-05-22 src/main.cpp

@@ -2,7 +2,7 @@
 * @Author: UnsignedByte
 * @Date:   2021-04-15 13:21:00
 * @Last Modified by:   UnsignedByte
-* @Last Modified time: 2021-05-13 18:16:11
+* @Last Modified time: 2021-05-13 18:39:12
 */
 
 #include <SFML/Graphics.hpp>
@@ -35,6 +35,7 @@ int WIDTH = 800;
 int HEIGHT = 800;
 float speed = 0.05; // time speed
 int fps = 30; // fps
+int trail_fps = 100; // how often to add to trail queue
 int duration = 60; // duration in seconds
 float trail_length = 5; //# of seconds of trail to save (will round)
 int _r_trail_length;
@@ -290,7 +291,7 @@ int main(int argc, char **argv)
 	}
 
 	int trail_count = trail_ids.size();
-	_r_trail_length = trail_count*trail_length/DT*speed;
+	_r_trail_length = trail_count*trail_length*trail_fps*speed;
 
 	L = std::min(WIDTH, HEIGHT)/static_cast<double>(N)/2;
 
@@ -308,7 +309,7 @@ int main(int argc, char **argv)
 	pendulums[0] = sf::Vector2f(WIDTH/2,HEIGHT/2);
 	char* filename;
 
-	asprintf(&filename, "output/%dPendulum%.2lfkg%.3lftheta_%dx%d_%dsecs%dfps_%.3fspeed%lfdt.mp4", N, M, ANGLE, WIDTH, HEIGHT, duration, fps, speed, DT);
+	asprintf(&filename, "output/%dPendulum%.2lfkg%.3lftheta_%dx%d_%dsecs%dfps_%.3fspeed%lfdt_%dtrails.mp4", N, M, ANGLE, WIDTH, HEIGHT, duration, fps, speed, DT, trail_count);
 
 	AVFormatContext* av_format_context;
 	AVCodecContext* av_codec_context;
@@ -410,6 +411,7 @@ int main(int argc, char **argv)
 	res = av_frame_get_buffer(frame, 32);
 
 	double lastFrame = 0;
+	double lastTrailFrame = 0;
 	sf::VertexArray trails_vertex(sf::Points, 0);
 
 	for (int i = 0; i < duration*speed/DT; i++) { //encode 60 seconds of video
@@ -438,15 +440,17 @@ int main(int argc, char **argv)
 
 			std::list<sf::Vector2f>::iterator it;
[…]
